\pkg{openTSNE}: A Modular \proglang{Python} Library for t-SNE Dimensionality Reduction and Embedding

### 2023-08-24 2023-08-final-revision/response-to-reviewers.tex

\vspace{5mm}

\noindent Dear Editor and the Reviewers, \\

\vspace{3mm}

\noindent Thank you for your valuable feedback for the final version of the manuscript. Below, we provide detailed, point-by-point responses to minor comments and change requests. To facilitate the review process, we also include a version of the manuscript with the highlighted changes from the previous revision.

@@

\noindent\textbf{Manuscript style comments}
\reviewc{r1q1}{Code should have enough spaces to facilitate reading.  Please include spaces before and after operators and after commas (unless spaces have syntactical meaning).}

\answer{As previously stated, when formatting the code used in the manuscript, we followed the standard Python code-style guidelines specified in PEP 8 (https://peps.python.org/pep-0008/). The style guide specifies that spaces should sometimes be omitted around operators, particularily when specifying parameter values. This is different from some other popular programming languages, such as R, but is considered standard practice in the Python community.}

\reviewc{r1q2}{As a reminder, please make sure that \textbackslash proglang,  \textbackslash pkg and \textbackslash code have been used for highlighting throughout the paper (including titles and references), except where explicitly escaped.}

\answer{We have inspected the manuscript to make sure that all package names, programming language references, and code snippets are surrounded by the appropriate Latex tags.}

@@

\reviewc{r3q1}{Page 5 mentions that Belkina et al suggest learning rate N/12. However, I checked the defaults in openTSNE 1.0 and the default learning rate seems to be N. Would be good to comment on that.}

\answer{
We have added a clarifying sentence to Page 5, and describe the default behaviour of openTSNE in the Implementation section in point 3 on page 8.
}

\reviewc{r3q2}{Page 21 says you used scikit-learn 1.1 for the benchmark, page 27 gives version 1.0. Also, given that there is version 1.2 out and there were some differences to t-SNE implementation in it, I wonder if it makes sense to use 1.2 for benchmarking (not sure if anything is going to change or not).}

\answer{
Thank you for pointing out this oversight, we have corrected the scikit-learn version to 1.1.2 in the caption of Table 1. We have also run several benchmarks comparing the performance of scikit-learn v1.1 vs v1.2 and the newest v1.3. We found no discernible differences in runtime, therefore, we feel that updating the manuscript in any large way is unnecessary. We have, however, added a corresponding footnote on page 20.
}

### 2023-08-24 2023-08-final-revision/main.tex
minor changes

A key component of the affinity model construction process is finding a specified number of nearest neighbors for each data point. \opentsne provides several backends for finding the nearest neighbors, each with its own advantages and drawbacks. For exact nearest neighbor search, we use the implementation from \pkg{scikit-learn}~\citep{pedregosa2011scikit}. While asymptotically slower than approximate nearest-neighbor search methods, exact approaches introduce minimal computation overhead and tend to run faster for smaller data sets. The \pkg{annoy}~\citep{annoylib} library offers a fast and efficient implementation of random-projection trees and supports several commonly used metrics but lacks support for sparse matrices. The \pkg{pynndescent}~\citep{pynndescent} library supports sparse matrices and allows users to define and use custom metrics. Finally, we include the \pkg{hnswlib}~\citep{hnswlib} library as it provides a third alternative algorithm for approximate nearest neighbor search, which may perform better on certain data sets.
By default, \opentsne automatically selects the most appropriate algorithm for a given data set based on data set size, sparsity, and metric support. These can be overridden by the user and easily extended to support custom nearest-neighbor search solutions.

@@

We adopt the learning rate $\eta=N/\rho$ proposed by \citet{belkina2019automated}, who suggest using a global learning rate computed from the early exaggeration factor $\rho$. Empirically, we have found that adapting the learning rate to the exaggeration factor $\rho$ used in the second phase of the optimization as well further speeds up convergence.

>>>>

A key component of the affinity model construction process is finding a specified number of nearest neighbors for each data point. \opentsne provides several backends for finding the nearest neighbors, each with its own advantages and drawbacks. We use the implementation from \pkg{scikit-learn}~\citep{pedregosa2011scikit} for the exact nearest neighbor search. While asymptotically slower than approximate nearest-neighbor search methods, exact approaches introduce minimal computation overhead and tend to run faster for smaller data sets. The \pkg{annoy}~\citep{annoylib} library offers a fast and efficient implementation of random-projection trees and supports several commonly used metrics but lacks support for sparse matrices. The \pkg{pynndescent}~\citep{pynndescent} library supports sparse matrices and allows users to define and use custom metrics. Finally, we include the \pkg{hnswlib}~\citep{hnswlib} library as it provides a third alternative algorithm for approximate nearest neighbor search, which may perform better on specific data sets.
By default, \opentsne automatically selects the most appropriate algorithm for a given data set based on data set size, sparsity, and metric support. These can be overridden by the user and easily extended to support custom nearest-neighbor search solutions.

@@

We adopt the learning rate $\eta=N/\rho$ proposed by \citet{belkina2019automated}, who suggest using a global learning rate computed from the early exaggeration factor $\rho$. Empirically, we have found that adapting the learning rate to the exaggeration factor $\rho$ used in the second phase of the optimization also further speeds up convergence.
